Versatile Robust Clustering of Ad Hoc Cognitive Radio Network

### 2016-09-04 journal-draft/bare_jrnl_comsoc.tex
modified a lot, to modify the game part

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Licensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, as one way, unlicensed users can access the spectrum via opportunistic spectrum access, \ie they access the licensed spectrum only after validating the channel is unoccupied by licensed users, where spectrum sensing~\cite{sensing_survey_2009} plays an important role in this process.
%
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), while the unlicensed users are referred to as secondary users and constitute the cognitive radio networks (CRN).
%
%cluster
Regarding to the operation of CRN, efficient spectrum sensing is identified to be critical for a smooth operation of a cognitive radio network~\cite{Sahai_FundamentalDesignTradeoffs2006}.
This can be achieved by cooperative spectrum sensing of multiple secondary users, which has been shown to cope effectively with noise uncertainty and channel fading, thus remarkably improving the sensing accuracy~\cite{coorperativeSensing_Akyildiz11}.
Collaborative sensing relies on the consensus of CR users\footnote{The terms user and node appear interchangeably in this paper. In particular, user is adopted when its networking or cognitive ability are discussed or stressed, while we refer node typicallly in the context of the topology.} within a certain area, in this regard, clustering is regarded as an effective method to realize cooperative spectrum sensing~\cite{Sun07_clustering_spectrum_secsing, Zhao07}.
% ... decreases considerably the negative probability due to the random channel effects like fading and shadowing.
%
Clustering is also efficient to coordinate the channel switch operation when primary users are detected by at least one CR node residing in the cluster.
The cluster head (CH) can enable all the CR devices within the same cluster to stop payload transmission swiftly on the operating channel and to vacate the channel ~\cite{willkomm08}.
%
%As to the coexistence of CR users, they can be notified by cluster head (CH) or other cluster members about the possible collision, the possibility for them to interfere neighbouring clusters is reduced~\cite{centralizedSharing80222}. 
In addition to the collaborate sensing advantage, the use of clusters is beneficial as it reduces the interference between cognitive clusters~\cite{centralizedSharing80222}.
Clustering algorithm has also been proposed to support routing in cognitive radio networks ~\cite{Abbasi_survey_07}.

@@

\section{Robust Clustering Problem in CRN}
\label{problem}
In this section, we describe the definition of cluster in the context of CRNs, and the problem of robust clustering.

@@

\begin{table}[h!]
\caption{Notations in robust clustering problem}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster $C$\\
%$\text{CH}$ & cluster head\\
$\delta$ & desired cluster size\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & social connectivity degree of CR node $i$\\
$C_i$  & the $i$th legitimate cluster (only appear in Sec.~\ref{centralized_opt})\\
\bottomrule
\end{tabular}
\end{table}

@@

In the introduction section, we have stated that cluster size should be given consideration to justify the concept of robustness of clusters, \ie without specifying requirement on cluster sizes, small clusters will be generated to obtain more CCCs.
Except for cooperative sensing, clusters need to conduct some other functionalities.
When cluster size is large, there will be substantial burden on cluster heads to manage the cluster members, which is a challenge for resource limited cluster heads, thus the cluster size should fall in a desired range~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.

The centralized clustering scheme aims to form clusters with desired sizes, meanwhile the total number of CCCs of all clusters is maximized.
In the following, we refer this problem as \textit{centralized clustering}, and give the formal problem definition.

@@

Given a cognitive radio network $\mathcal{N}$ where nodes are indexed from 1 to $N$ sequentially.
Based on certain correlation, certain secondary users constitute one cluster $C$.
$1\leq |C| \leqslant k$ where $|C|$ is the size of cluster $C$ and $k$ is a positive integer.
We name the collection of such clusters as $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$, where $\mathcal{S}$ satisfies the following properties: $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$ and $K(C(i))\neq \emptyset$ for any $i$ which satisfies $1\leq i \leq |\mathcal{S}|$.

%, which distinguishes the centralized clustering problem discussed in this paper.
We give a new definition of the number of CCCs, where the number of common control channels is $|K(C)|$ if $|C|>1$, and is zero when $|C|=1$.
We use $f(C)$ to denote the number of CCCs of a cluster $C$ in the new definition.

The centralized clustering problem is to find a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, so that $\bigcup_{C_j\in \mathcal{S}'} C_j = N$, and $C_{j'}\cap C_j =\emptyset$ for $C_{j'}, C_j\in \mathcal{S}'$ and $j'\neq j$, so that $\sum_{C\in \mathcal{S}'} f(C)$ is maximized.
% when $|C|>1$, and $f(\cdot)=0$ when $|C|=1$.
%As to $f(\cdot):z\rightarrow z$, under this problem setting, if only number of common channels are given consideration, only singleton clusters will be preferred, which contradicts to our goal of clustering CR nodes together, thus we choose $f(\cdot)= |K_{C}|\cdot |C|$, the product of number of common channels and cluster size.
The decision version of centralized clustering in CRN is to ask whether there exists a no-empty $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f \geqslant \lambda$ where $\lambda$ is a real number.% in stead of to maximize $\sum_{C\in \mathcal{S}'} f$.
\end{mydef}

@@

\subsection{Complexity of the Robust Clustering Problem}
In this section we investigate the complexity of robust clustering problem.
Theorem~\ref{theorem1} tells centralized clustering problem in CRN is one NP-hard problem.

\begin{theorem}
\label{theorem1}
CRN clustering problem is NP-hard, when the maximum size of clusters $k\geqslant 3$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
In the following sections, both centralized and distributed solutions will be introduced.

\section{Centralized Solution for Robust Clustering}
\label{centralized_opt}
%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
As the robust clustering problem in CRN is NP hard, there is no polynomial algorithm to solve the problem, \ie maximizing the sum of CCCs in all non-singleton clusters.
Thus, we formulate the problem into a binary linear programming problem, where the objective function and the constraints are heuristic, in particular, the objective is different from the object in the Definition~\ref{def_centralized_clustering}.
%Note that binary linear programming is in NP-complete.

@@

Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{legitimate} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
Legitimate clusters are the clusters which satisfy the conditions in Section~\ref{def_cluster}. 
Note that the legitimate clusters include the \textit{singleton} ones, \ie the cluster which has only one CR node, so that we can guarantee the partition of any network is always feasible.
%
With $N=|\mathcal{N}|, G=|\mathcal{G}|$, we construct a constant $G\times N$ matrix $Q_{G\text{x}N}$. 
The element of matrix $Q$ is $q_{ij}$, where the subscript $i$ is the index of legitimate cluster, and $j$ is the node ID of one CR node.
%Note that $C_i$ means $i$th cluster in the collection $\mathcal{G}$, doesn't denote the cluster where cluster head is $i$, this notation is only valid in this subsection.
There are $i\in \{1,2,\cdots,G-1, G\}$, and $j\in \{1,2,\cdots,N-1, N\}$.
Element $q_{ij}= |K(C_i)|$ if node $j\in C_i$, and $q_{ij}= 0$ if $j\notin C_i$.
In other words, each non-zero element $q_{ij}$ denotes the number of CCCs of the cluster $i$ where node $j$ resides.

\begin{figure}[ht!]
\centering
\bordermatrix{~ 		& 1 	& 2 	& 3 	& \cdots & j & \cdots	& N-1 	& N	\cr
                  1 	& |K(C_1)| 	& |K(C_1)| 	& 0 	& \cdots & \cdots &\cdots	& 0 	& 0	\cr
                  2 	& |K(C_2)| 	& 0 	& |K(C_2)| 	& \cdots & \cdots & \cdots 	& 0 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots		& 		& \vdots \cr
				i 	& 0 	& |K(C_i)| 	& 0 	& \cdots  & \cdots & \cdots 	& |K(C_i)| 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots & \cdots & \cdots 		& 		& \vdots \cr
				\vdots 	& \vdots  	& 0 	& 0 	& \cdots & \cdots & \cdots 	& |K(C_i')| 	& 0	\cr
				G  	& |K(C_G)| & \cdots	 	& 		&  \vdots	& \cdots & \cdots& 		& \vdots \cr}	
\caption{An example of Matrix $Q$, its rows correspond to all legitimate clusters, and columns correspond to the CR nodes in the CRN.}
\label{xx}
\end{figure}

We build a $G\times N$ binary variable matrix $X$, which illustrates the clustering solution.
The element of matrix $X$ is binary variable $x_{ij}, i=1, \ldots, G, j=1, \ldots, N$.
%$x_{ij}=1$ denotes node $i$ is in the $j$th legitimate cluster, and the cluster is chosen in the solution, $x_{ij}=0$ means the $j$th legitimate cluster is not adopted by the partition.
%Note that matrix $Q$ contains only constant elements, and matrix $X$ contains only binary variables.
Now, we can formulate the optimization problem as follows,

@@

The objective is a sum of two components, the first component is the sum of products of cluster size and the corresponding number of CCCs, which is the sole metric adopted by the scheme SOC~\cite{LIU_TMC11_2}.
The second component is the \textit{punishment} for choosing the clusters whose sizes are not $\delta$.
In fact, the second part is particularly designed to eliminates the drawbacks of SOC, \ie SOC produces a large number of singleton clusters and a few large clusters.
In practical computation, we minimize the opposite of the first part, and the punishment is a positive value.
%
The first constraint restricts each node $j$ to reside in exactly one cluster.
In the second constraint, $w_i$ is an auxiliary binary variable which denotes whether cluster $C_i$ is chosen by the solution, in particular, 
$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th legitimate cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th legitimate cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
The second constraint regulates that when the $i$th legitimate cluster $C_i$ is chosen, the number of elements which equal to 1 in the $i$th row of the matrix $X$ is $|C_i|$.
%
Now we explain how does the mechanism of the punishment in the objective work. 
The parameter $p$ is defined as follows,
$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$} \\
\dots
\end{array} \right.
$$
where $\alpha_i>0$ and increases when $|C_i|$ diverges from $\delta$.
Because of $w_i$, any chosen cluster ($w=0$) brings certain \textit{punishment}.
When the chosen cluster's size is desired size $\delta$, the punishment is zero.
In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
In particular, when $w_i=0$ and $|C_i|=1$, the punishment is the most severe.
This design doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
Choice of $\alpha_i$ affects the resultant clusters.

@@

%Note that the proposed centralized solution is heuristic.
%We reiterate the reasons for pursuing the heuristic scheme, 
%Note the collection of legitimate clusters is dependant on the network topology and spectrum availability in the network, thus to each specific CRN, the space of solution is different.

\section{Distributed Clustering Algorithm: ROSS}
\label{ross}
The cluster formation of one CRN can be obtained by solving the proposed optimization problem, but centralized scheme is inherently not suitable for the CRN which involves dynamics, \eg the mobility of CR users and the variance of the available spectrum due to primary users' operation.
In this section we introduce the distributed clustering scheme ROSS.
%
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters based on the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

\subsection{Phase I - Cluster Formation}
\label{phaseI}
After conducting spectrum sensing and communication with neighbours, every CR node is aware of the available channels available for themselves as well as for all its neighbors.
For each CR user $i$, two metrics are proposed to characterize the spectrum proximity between $u$ and its neighborhood:
%\newtheorem{def1}{Definition}
%\label{def1}
%\begin{def1}
%Connection vector $\left\{d_i,g_i\right\}$ of CR node $i$:
%spectrum connectivity degree: 
%\begin{equation}
%\label{d_i}
%d_i=\sum_{j\in N_i}\vert V_i\cap V_j\vert
%\end{equation}
%Social Connection degree:
%\begin{equation}
%\label{g_i}
%g_i=|\bigcap_{j\in N_i}V_j|
%\end{equation}
%\end{def1}
\begin{itemize}

\item \textit{Individual connectivity degree} $d_i$: $d_i=\sum_{j\in \text{Nb}(i)}\vert K_i\cap K_j\vert$. It denotes the sum of the numbers of common controls channels between node $i$ and every neighbour.
It is an indicator of node $i$'s adhesive property to the CRN.

\item \textit{Social connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. It is the number of common channels available to $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbours.
\end{itemize}
Individual connectivity degree $d_i$ and social connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where each node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	% The number of common channels between two nodes is shown by the edge.
	\label{fig1}
\end{figure}

After introducing the connectivity vector, we proceed to introduce the first phase of algorithm ROSS.
To put it briefly, in this phase cluster heads are determined in the beginning. Then clusters are formed on the basis of the cluster heads' neighborhoods.

>>>>

\section{Introduction}
\label{intro}
%CR concept
\IEEEPARstart{C}{ognitive} radio (CR) is a promising technology to solve the spectrum scarcity problem~\cite{Mitola}.
Licensed users access the spectrum allocated to them whenever there is information to be transmitted.
In contrast, as one way, unlicensed users can access the spectrum via opportunistic spectrum access, \ie they access the licensed spectrum only after validating the channel is unoccupied by licensed users, where spectrum sensing~\cite{sensing_survey_2009} plays an important role in this process.
%
In this hierarchical spectrum access model~\cite{zhao_survey_DSA_2007}, the licensed users are also called primary users (PU), while the unlicensed users are referred to as secondary users and constitute the cognitive radio networks (CRN).
%
%cluster
Regarding to the operation of CRN, efficient spectrum sensing is identified to be critical for a smooth operation of a cognitive radio network~\cite{Sahai_FundamentalDesignTradeoffs2006}.
This can be achieved by cooperative spectrum sensing of multiple secondary users, which has been shown to cope effectively with noise uncertainty and channel fading, thus remarkably improving the sensing accuracy~\cite{coorperativeSensing_Akyildiz11}.
Collaborative sensing relies on the consensus of CR users\footnote{The terms user and node appear interchangeably in this paper. In particular, user is adopted when its networking or cognitive ability are discussed or stressed, while we refer node typicallly in the context of the topology.} within a certain area, in this regard, clustering is regarded as an effective method to realize cooperative spectrum sensing~\cite{Sun07_clustering_spectrum_secsing, Zhao07}.
% ... decreases considerably the negative probability due to the random channel effects like fading and shadowing.
Clustering is a process of grouping certain users in a proximity into a collective.
Clustering is also efficient to coordinate the channel switch operation when primary users are detected by at least one CR node residing in the cluster.
The cluster head (CH) can enable all the CR devices within the same cluster to stop payload transmission swiftly on the operating channel and to vacate the channel ~\cite{willkomm08}.
%
%As to the coexistence of CR users, they can be notified by cluster head (CH) or other cluster members about the possible collision, the possibility for them to interfere neighbouring clusters is reduced~\cite{centralizedSharing80222}. 
In addition to the collaborate sensing advantage, the use of clusters is beneficial as it reduces the interference between cognitive clusters~\cite{centralizedSharing80222}.
Clustering algorithm has also been proposed to support routing in cognitive radio networks ~\cite{Abbasi_survey_07}.

@@

\section{Robust Clustering Problem in CRN}
\label{problem}
In this section, we present the definition of cluster in the context of CRNs, define the clustering problem and discuss its complexity.

@@

\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & collection of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h_C$ & the cluster head of a cluster C\\
$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$C_i$  & the $i$th potential cluster (only used in Sec.~\ref{centralized_opt})\\
\bottomrule
\end{tabular}
\end{table}

@@

The clustering scheme we propose in the paper aims to form clusters with preferred sizes, meanwhile maximize the average number of CCCs per cluster.
%In the following, we refer this problem as \textit{centralized clustering}. 
%We have mentioned that the sizes of the formed clusters should be given consideration, which is import for the discussion on cluster robustness, \ie 
when perusing the cluster robustness without specifying the requirement on cluster sizes, a clustering method will be inclined to form small clusters.
In that case, although more CCCs are obtained per cluster, the benefits brought in by the collective of the cluster members are compromised.
%Except for cooperative sensing, clusters need to conduct some other functionalities.
Large clusters are not preferred in some scenarios neither, \eg for the CRN composed with resource limited users, managing the cluster members is a substantial burden. 
Hence, the cluster size should fall in a desired range~\cite{Chen04clusteringalgorithms, capacity_cluster_06}.
In the following, we present the formal definition of robust clustering problem.

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$, secondary users are indexed from 1 to $N$.
A cluster $C$ can be composed with one or multiple secondary users when $|K(C)|>0$, and any two cluster members can communicate either directly, or through other members belonging to the same cluster.
We use $f(C)$ to denote the number of CCCs of a cluster $C$, and give it a new definition as follows, the number of common control channels is $f(C) = |K(C)|$ if $|C|>1$, and $f(C)=0$ when $|C|=1$.
The size of cluster $C$ is not bigger than a positive integer $k$.
The collection of such clusters $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$, satisfies the following condition: $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$ and $1\leq i \leq |\mathcal{S}|$.

%, which distinguishes the centralized clustering problem discussed in this paper.

The robust clustering problem is to find a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f(C)$ is maximized, and $\mathcal{S}'$ satisfies the following conditions, $\bigcup_{C_j\in \mathcal{S}'} C_j = \mathcal{N}$ and $C_{j'}\cap C_j =\emptyset$ where $C_{j'}, C_j\in \mathcal{S}'$ and $j'\neq j$.
% when $|C|>1$, and $f(\cdot)=0$ when $|C|=1$.
%As to $f(\cdot):z\rightarrow z$, under this problem setting, if only number of common channels are given consideration, only singleton clusters will be preferred, which contradicts to our goal of clustering CR nodes together, thus we choose $f(\cdot)= |K_{C}|\cdot |C|$, the product of number of common channels and cluster size.
\end{mydef}

@@

\subsection{Complexity of the Robust Clustering Problem}
%In this section we investigate the complexity of robust clustering problem.
Based on Definition~\ref{def_centralized_clustering}, the decision version of this problem is to ask whether there exists a no-empty $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f(C) \geqslant \lambda$ where $\lambda$ is a real number.
We have the following theorem on complexity.

\begin{theorem}
\label{theorem1}
Robust clustering problem in CRN is NP-hard, when the maximum size of clusters $k\geqslant 3$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.
As it is not likely to generate clusters with size of 2, this theorem actually indicates that the robust clustering problem in CRN is NP-hard. 
In the following sections, both centralized and distributed solutions will be introduced.

\section{Centralized Solution for Robust Clustering}
\label{centralized_opt}
%Exact cover problem can be solved with Knuth's Algorithm X~\cite{dancingLinks_Knuth} as it finds out all the instances of exact cover, then we can choose the one with the biggest sum of weights. 
As the robust clustering problem in CRN is NP hard, there is no polynomial solution available.
In this paper we formulate the problem into a binary linear programming problem.
To make the formulation possible, we adopt heuristics in both the objective function and the constraints, besides, the objective is not totally aligned with the object in the Definition~\ref{def_centralized_clustering}.

@@

@@

The objective is to maximize the sum of the products of cluster size and the corresponding number of CCCs.
The second component is designed to be the \textit{punishment} for forming clusters whose sizes are not $\delta$, which is the preferred cluster size.
We minimize the opposite of the real objective to make this problem solvable.
The constraints guarantee to obtain the clusters which don't overlap.
Now we introduce the formulation in details.

$N=|\mathcal{N}|, G=|\mathcal{G}|$.
$\mathcal{G}$ is the collective of all the \textit{potential} clusters in a CRN $\mathcal{N}$, and the sizes of these clusters are $1,2,\ldots,\delta$, where $\delta$ is the preferred size for clusters.
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
Potential clusters are the clusters which satisfy the conditions in Section~\ref{def_cluster}. 
Note that the potential clusters include the \textit{singleton} ones, \ie the cluster which has only one CR node.
By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.

Constant $q_{ij}$ is the element of the constant $G\times N$ matrix $Q_{G\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
The subscript $i$ is the index of a potential cluster, $j$ is the node ID of a CR node, $q_{ij}= |K(C_i)|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
In other words, each non-zero element $q_{ij}$ denotes the number of CCCs of the cluster $i$ where node $j$ resides.
Knowing the meanings of subscripts $i$ and $j$, it is easy to notice there are $i\in \{1,2,\cdots,G-1, G\}$ and $j\in \{1,2,\cdots,N-1, N\}$.
%
The binary variable $x_{ij}$'s subscripts have the same meanings with that of $q_{ij}$, thus they have the same scope.
$x_{ij}$ indicates whether the CR node $j$ resides in the potential cluster $i$, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
%

\begin{figure}[ht!]
\centering
\bordermatrix{~ 		& 1 	& 2 	& 3 	& \cdots & j & \cdots	& N-1 	& N	\cr
                  1 	& |K(C_1)| 	& |K(C_1)| 	& 0 	& \cdots & \cdots &\cdots	& 0 	& 0	\cr
                  2 	& |K(C_2)| 	& 0 	& |K(C_2)| 	& \cdots & \cdots & \cdots 	& 0 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots		& 		& \vdots \cr
				i 	& 0 	& |K(C_i)| 	& 0 	& \cdots  & \cdots & \cdots 	& |K(C_i)| 	& 0	\cr
				\vdots  	&\vdots & 	 	& 		&  \vdots & \cdots & \vdots 		& 		& \vdots \cr
				\vdots 	& 0  	& 0 	& 0 	& \cdots & \cdots & \cdots 	& |K(C_i')| 	& 0	\cr
				G  	& |K(C_G)| & \cdots	 	& 		&  \vdots	& \cdots & \vdots& 		& \vdots \cr}	
\caption{An example of Matrix $Q$, its rows correspond to all potential clusters, and columns correspond to the CR nodes in the CRN. }
\label{costant_matrix_Q}
\end{figure}

Now let us have a look at the objective function again.
Although the related work SOC~\cite{LIU_TMC11_2} adopts the product of cluster size and number of CCCs as its sole metric, SOC still generates a large number of singleton clusters and a few large clusters.
To refrain the clusters which don't have preferred clusters sizes, the second item is designed, where

$$
w_i = \left\{ \begin{array}{rl}
0 &\mbox{if $i$th potential cluster $C_i$ is chosen} \\
1 &\mbox{if $i$th potential cluster $C_i$ is not chosen} \\
\end{array} \right.
$$
$w_i$ is an auxiliary binary variable which denotes whether cluster $C_i$ is chosen by the clustering solution or not.
When $w_i$ is zero, it means $C_i$ is chosen by the clustering solution, in this case, the second item in the objective function is $p$, which is defined as follows,

$$
p = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\alpha_1 &\mbox{if $|C_i|=\delta-1$} \\
\alpha_2 &\mbox{if $|C_i|=\delta-2$} \\
\dots
\end{array} \right.
$$
where $\alpha_x$ is a positive value, which increases with the divergence between $|C_i|$ and $\delta$, thus these is $\alpha_2> \alpha_1>0$.
Because of $w_i$, any cluster which appears in the clustering solution ($w=0$) will bring certain \textit{punishment}.
Only when that cluster's size is exactly the preferred size $\delta$, the punishment is zero.
In contrary, when the chosen cluster's size diverges from $\delta$, the objective function suffers \textit{loss}.
In particular, when $w_i=0$ and $|C_i|=1$, the punishment is the most severe.
With such design, the solution is encouraged not to generate singleton clusters.
%doesn't follow the definition of $f(C)$ in Definition~\ref{def_centralized_clustering} strictly, where $f(C_i)=0$ when $|C_i|=1$, but our design echoes the definition by exerting the most severe punishment on the singleton clusters in the clustering solution.
%Choice of $\alpha_i$ affects the resultant clusters.

As to the constraints, the first constraint restricts each node $j$ to reside in exactly one cluster.
The second constraint regulates that when the $i$th potential cluster $C_i$ is chosen, there will be exactly $|C_i|$ CR nodes residing in cluster $C_i$.

@@

%Note that the proposed centralized solution is heuristic.
%We reiterate the reasons for pursuing the heuristic scheme, 
%Note the collection of potential clusters is dependant on the network topology and spectrum availability in the network, thus to each specific CRN, the space of solution is different.

\section{Distributed Clustering Algorithm: ROSS}
\label{ross}
The centralized robust clustering scheme is inherently not suitable for the CRN which involves dynamics in multiple aspects.
Whenever clustering or re-clustering is necessary, which could be triggered by either the mobility of CR users or the variance of the available spectrum, the centralized entity needs to collect the information from all the CR nodes, and then distributes the clustering solution after computation.

In this section we introduce the distributed clustering scheme ROSS.
ROSS consists of two cascaded phases: \textit{cluster formation} and \textit{membership clarification}.
With ROSS, CR nodes form clusters based on the proximity of the available spectrum in their neighbourhood. Afterwards, CR nodes belong to one certain cluster. %by interacting with certain neighbours.
%is based on the spectrum sensing results in its neighbours , then

\subsection{Phase I - Cluster Formation}
\label{phaseI}
After conducting spectrum sensing and communicating with neighbors, every CR node is aware of the channels which are available on themselves and also those on their one hop neighbors.
Two metrics are proposed to characterize the proximity in terms of available spectrum between CR node $i$ and its neighborhood:

\item \textit{Individual connectivity degree} $d_i$: $d_i=\sum_{j\in \text{Nb}(i)}\vert K_i\cap K_j\vert$. 
$d_i$ is the total number of the CCCs between node $i$ and every its neighbor.
It is an indicator of node $i$'s adhesion to the CRN.

\item \textit{neighborhood connectivity degree} $g_i$: $g_i=|\bigcap_{j\in \text{Nb}(i)\cup i}K_j|$. 
It is the number of CCCs which are available for $i$ and all its neighbors.
$g_i$ represents the ability of $i$ to form a robust cluster with its neighbors.
\end{itemize}
Individual connectivity degree $d_i$ and neighborhood connectivity degree $g_i$ together form the \textit{connectivity vector}.
Figure~\ref{fig1} illustrates an example CRN where every node's connectivity vector is shown.	
\begin{figure}[ht!]
  \centering
\includegraphics[width=0.7\linewidth]{figure1.pdf}
% \includegraphics{figure1.pdf}
	\caption{Connectivity graph of the example CRN and the connectivity vector $(d_i, g_i)$ for each node. The sets of the indices of the available channels sensed by each node are: $K_A=\{1,2,3,4,5,6,10\}, K_B=\{1,2,3,5,7\}, K_C=\{1,3,4,10\}, K_D=\{1,2,3,5\}, K_E=\{2,3,5,7\}, K_F=\{2,4,5,6,7\}, K_G=\{1,2,3,4,8\}, K_H=\{1,2,5,8\}$. Dashed edge indicates the end nodes are within each other's transmission range.}
	\label{fig1}
\end{figure}

After introducing the connectivity vector, we proceed to introduce the first phase of algorithm ROSS.
In this phase, cluster heads are determined through a series of autonomous negotiations among the CR users, then the neighborhoods of the cluster heads become clusters.

\subsubsection{Determining Cluster Heads and Form Clusters}
In this phase, each CR node decides whether it is a cluster head by comparing its connectivity vector with its neighbors.
When CR node $i$ has lower individual connectivity degree than all its neighbors except for those which have already identified to be cluster heads, node $i$ becomes a cluster head.
%In other words, CR node $i$ becomes cluster head if $d_i > d_k, \forall k\in \texttt{Nb}_i\setminus CHs$ ($CHs$ donate the cluster heads existing in $\texttt{Nb}_i$).
If there is another CR node $j$ in its neighborhood which has the same individual connectivity degree as $i$, \ie $d_j = d_i$ and $d_j < d_{k}, \forall k\in \text{Nb}(j)\setminus \{\text{CHs}\cup i\}$, then the node between $i$ and $j$, which has higher neighborhood connectivity degree will become the cluster head, and the other node become one member of the newly identified cluster head. 
If $g_i = g_j$ as well, the node ID is used to break the tie, \ie the one with smaller node ID becomes a cluster head.
%
The node which is identified as a cluster head broadcasts a message to notify its neighbors of this change, and its neighbors which are not cluster heads become cluster members
\footnote{The reasons for the occurrence of the cluster heads in the neighborhood of a new cluster head will be explained in Section~\ref{ross_p1_guarantee_ccc} and \ref{ross_p2_cluster_pruning})}.
The pseudo code for the cluster head decision and the initial cluster formation is shown in Algorithm~\ref{alg0} in the appendix.
